---
title: "Lab 2 - Statistics and Data Analysis II"
output: 
  pdf_document:
    latex_engine: xelatex
author: "Zhen Liu"
date: "November 20, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
This lab is done by Zhen Liu

```{r Task 1}
data <- read.csv("hsbc_lab2.csv")

str(data)
summary(data)

model1 <- lm(lifesat ~ BMI, data = data)
summary(model1)
```
The coefficient represents the change in life satisfaction for one unit increase in BMI. The p-value tests the null hypothesis that β = 0 means BMI has no effect on life satisfaction. Small p-value suggests that BMI significantly influences life satisfaction.

```{r Task 2}
model_quad <- lm(lifesat ~ BMI + I(BMI^2), data = data)

summary(model_quad)
```
The coefficient for BMI^2 determines whether the relationship is nonlinear. If the coefficient is negative, the curve opens downward, suggesting a peak in life satisfaction at a certain BMI value. If positive, the curve opens upward.

```{r Task 3}
model_interaction <- lm(lifesat ~ BMI * sex, data = data)

summary(model_interaction)
```
The β1 shows the effect of BMI on life satisfaction for the reference gender。
The β2 shows the difference in life satisfaction between genders when BMI = 0.
The β3 indicates how the effect of BMI on life satisfaction differs between boys and girls.


```{r Task 4}
hypo_girls <- data.frame(sex = "Girl", BMI = c(20, 30))

hypo_girls$lifesat <- predict(model_interaction, newdata = hypo_girls)

print(hypo_girls)
```



```{r Task 5}
hypo_boys <- data.frame(sex = "Boy", BMI = c(20, 30))

hypo_boys$lifesat <- predict(model_interaction, newdata = hypo_boys)

print(hypo_boys)
```

```{r Task 6}
all_predictions <- rbind(hypo_girls, hypo_boys)

library(ggplot2)
ggplot(all_predictions, aes(x = BMI, y = lifesat, color = sex, group = sex)) +
  geom_line() +
  geom_point(size = 3) +
  labs(
    title = "Predicted Life Satisfaction by BMI and Gender",
    x = "BMI",
    y = "Life Satisfaction",
    color = "Gender"
  ) +
  theme_minimal()
```

```{r Task 7}
plot(model_interaction$fitted.values, residuals(model_interaction),
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")

hist(residuals(model_interaction), main = "Histogram of Residuals", xlab = "Residuals")

qqnorm(residuals(model_interaction))
qqline(residuals(model_interaction), col = "red")
```

7.a: the residuals appear randomly scattered around zero, the assumption of homoscedasticity is likely met.

7.b: If the residuals follow a straight line in the Q-Q plot and are approximately bell-shaped in the histogram, the normality assumption is likely met.

7.c: There might be other factors should be considered such as family income, education time, living quality and friendship etc. They can exert influence on the life satisfaction.

7.d: The data shows surveyed school children in 2014. Inferences from the analysis are limited to this population groups and may not fit for other demographics.